Inpainting with controlled dimensionality reduction in the Fourier domain

### 2011-06-21 project/Paper/paper.tex
Paper: Global parameter optimization completed.

  \path [line] (node7) |- (node11);
\end{tikzpicture}
\caption{Our newly proposed algorithm as a flowchart.}
\end{figure}

As outlined in the introduction, we combined several ideas for obtaining an inpainting algorithm that provides both good results and fast performance. The input of our algorithm is the image to be restored together with a mask that indicates its unknown spots.\\
Before entering the iterative part of our algorithm, the following three steps are performed (as seen in figure ??):
\begin{enumerate}
\item In order to guide the elimination of frequencies by finding the optimal threshold (as described below), we first remove additional pixels from the image and use them as the validation set. The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section ??).
\item Since our reduction in the Fourier space works best with framed patches, we frame the complete image by mirroring its margin. (The size of the frame is another optimizable global parameter.)
\item For obtaining an initial reconstruction of the image, we compute a weighted sum over all the neighboring pixels that are known. By setting all unknown pixels in the image to zero and encoding known values with 1s in the mask, elementwise division of the filtered image by the similarly filtered mask results in a normalized sum weighted by the filtering kernel. (This procedure, which we call Gaussian interpolation, allows us to use MATLAB's fast native functions and to fill holes up to the kernel size, which is another global parameter.)
\end{enumerate}

@@

\subsection{Global parameter optimization}
As shown in table \ref{parameters}, there are a total of 10 parameters in our algorithm that can be set. For many of them, it is quite clear what values should be used. For example, the size of the gauss kernel should be a bit larger than the biggest assumed hole in the image. For others, like the maximum number of iterations, it is not so clear. And yet others have a clear speed / accuracy tradeoff, like the stopping conditions of the threshold optimization.\\
Because we thought it not feasable to tune all the parameters by hand, we wrote an algorithm to do the global parameter optimization. The idea is to do this once for a few representative images, and get parameters that can model as closely as possible many of the images we are interested in. Our proposed method combines a gradient descent method with ideas from cross-validation, and tunes our 10-dimensional parameter-space, minimizing time and maximizing accuracy.

\subsubsection{Error function}
When doing a minimization of any kind, an error function is needed. Of course we already had the error function of the evaluation script, that gives a mean over the squared differences of the reconstructed to the original image. However, using only an error measure, the time would potentially grow to infinity to get the best accuracy possible during tuning.\\
Therefore we tried different error function including both, the image error $E$, and the time $T$ spent (see table \ref{error_function_table}). The image error here is the average squared error of a reconstructed pixel to the original pixel, scaled with a factor (usually about 10000), while the time is measured in seconds.

\begin{table*}
  \centering
  \begin{tabular}{|l|p{5cm}|p{5cm}|}
    \hline
    Function&Advantages&Disadvantages\\
    \hline
    $\exp(E)+\exp(T)$ & Increases cost fast to avoid $E$ or $T$ over a certain threshold & Not enough gradient for very low $E$, $T$ values\\
    \hline
    $(1-\exp(E))*(1-\exp(T))$ & Makes it possible to have large times for very low errors & Works also the other way around, and optimization only minimizes time\\
    \hline
    $E + T$ & Everywhere a good gradient & Time from 60 to 59 seconds is same \textit{improvement} as time from 2 to 1 second\\
    \hline
    $\begin{cases}
  E  & \text{if }T<60\\
  E+\exp(T/60 - 1) - 1 & \text{otherwise}
\end{cases}$ & Optimizes for error only when time is low enough. Function still continous, penalizes high times. & None so far, this is the function used by us\\
    \hline
  \end{tabular}
  \caption{Different combined error functions}
  \label{error_function_table}
\end{table*}

% TODO maybe add 3d plot for some of the error functions

\subsubsection{Gradient descent}
The idea behind our gradient descent is to optimize all the parameters at once, instead of fine-tuning each of them separately. For that, we
calculate the derivative of the error function depending on one of our 10 parameters, and then move the parameter into the direction where the
error decreases fastest. We also consider the slope, and make big changes if the derivative is big, and small changes if the derivative is small.
Iteratively, we calculate all of the 10 derivatives, change all the parameters and start over again.

Since the error and time are measured, we cannot algebraically determine the gradient of the error function. Instead, we do a discrete estimation,
measuring the error with finite differences. The error for parameter values plus and minus a small delta, divided by two, gives us our estimated
derivative.

In figure \ref{}

@@

\subsection{Figures and Tables}

>>>>

  \path [line] (node7) |- (node11);
\end{tikzpicture}
\caption{Our newly proposed algorithm as a flowchart.}
\label{flowchart}
\end{figure}

As outlined in the introduction, we combined several ideas for obtaining an inpainting algorithm that provides both good results and fast performance. The input of our algorithm is the image to be restored together with a mask that indicates its unknown spots.\\
Before entering the iterative part of our algorithm, the following three steps are performed (as seen in figure \ref{flowchart}):
\begin{enumerate}
\item In order to guide the elimination of frequencies by finding the optimal threshold (as described below), we first remove additional pixels from the image and use them as the validation set. The fraction of additional pixels to be removed is a global parameter, which we optimized using gradient descent (as explained in section \ref{gradient_descent}).
\item Since our reduction in the Fourier space works best with framed patches, we frame the complete image by mirroring its margin. (The size of the frame is another optimizable global parameter.)
\item For obtaining an initial reconstruction of the image, we compute a weighted sum over all the neighboring pixels that are known. By setting all unknown pixels in the image to zero and encoding known values with 1s in the mask, elementwise division of the filtered image by the similarly filtered mask results in a normalized sum weighted by the filtering kernel. (This procedure, which we call Gaussian interpolation, allows us to use MATLAB's fast native functions and to fill holes up to the kernel size, which is another global parameter.)
\end{enumerate}

@@

\subsection{Global parameter optimization}
\subsubsection{Optimized parameters}
As shown in table \ref{parameters}, our algorithm has a total of 10 parameters that can be set globally. Instead of tuning them by hand, we automated the optimization process and implicitly use the best parameters found so far, which we store in the file system. This feature also allows us to interrupt the optimization process at any time and continue with the best parameters later on.

\begin{table*}
\begin{center}
\begin{tabular}{|l|l|r|r|}
\hline
Parameter & Description & Initial value & Optimized value\\
\hline
gauss\_size & Size of the Gaussian kernel & ? & ? \\
gauss\_sigma & Sigma of the Gaussian kernel & ? & ? \\
patch\_size & Patch size as exponent to the basis 2 & ? & ? \\
patch\_frame\_size & Frame size in pixels & ? & ? \\
td\_abortbelow\_stdev & Minimal standard deviation while evaluating thresholds & ? & ? \\
td\_abortbelow\_stepsize & Minimal step size while evaluating thresholds & ? & ? \\
td\_middle & Initial threshold & ? & ? \\
validation & Fraction of pixels in the validation set & ? & ? \\
max\_iterations & Maximum number of iterations & ? & ? \\
abortbelow\_change & Minimal change for performing another iteration & ? & ? \\
\hline
\end{tabular}
\end{center}
\caption{Parameters with their initial and optimized values.}
\label{parameters}
\end{table*}

\subsubsection{Gradient descent}
\label{gradient_descent}
We optimize all parameters at once using a method called gradient descent. Since our cost function is based on measurements, we cannot derive it with respect to individual parameters and have to determine the gradient with finite differences. Besides weighting the parameters with respect to their improvements (which gives us the slope), we also defined a momentum term that should help us to overcome local minima.

\subsubsection{Cost function}
For optimizing a set of parameters, a cost function is needed. Its most important component is obviously the error function used in the submission system, which calculates the mean over the squared differences between the reconstructed and the original image. However, without including the performance of the algorithm, the runtime could grow arbitrarily to get the best accuracy possible.\\
We considered several cost functions based on the error $E$ and the time $T$ (see table \ref{error_functions}). These two values are scaled appropriately: Usually, the error is multiplied by a factor of 10'000 and the time (measured in seconds) by a factor of 0.01.

\begin{table*}
\centering
\begin{tabular}{|l|p{6cm}|p{6cm}|}
\hline
Function & Advantages & Disadvantages \\
\hline
$E + T$ & Easy to understand and everywhere a good gradient & Linear improvement does not necessarily follow one's intuition about the ordering of results \\
\hline
$\exp(E)+\exp(T)$ & Fast increase for both error and time & Not enough gradient for very low $E$ and $T$ values \\
\hline
$(1-\exp(E))*(1-\exp(T))$ & Allows very large times for very low errors & Also true vice versa, i.e. a minimization in time \\
\hline
$\begin{cases}
E & \text{if }T<60\\
E+\exp(T/60 - 1) - 1 & \text{otherwise}
\end{cases}$ & Continuous function that optimizes for error only when the time is below a certain value but penalizes long times. & Seems to work quite well and therefore used by us \\
\hline
\end{tabular}
\caption{Various cost functions used in the optimization process.}
\label{error_functions}
\end{table*}

\subsubsection{Sample data}
We used the three images provided for the last exercise (consisting of both natural and artificial images) to evaluate the accuracy and performance of the chosen parameters. Instead of using the supplied mask, however, we generate one randomly for every evaluation in order to prevent overfitting to a particular mask.

\subsubsection{Found values}
Our observations during the optimization process can be split into three classes:
\begin{enumerate}
\item Some parameters converged towards their expected value. The size of the Gaussian kernel, for example, has to be slightly bigger than the largest hole assumed in an image.
\item Some parameters have a clear speed vs. accuracy tradeoff, like the stopping conditions of the threshold optimization, which depend on our cost function and could not be foreseen.
\item Some parameters assumed totally different values than anticipated: The patch size became much larger and the validation set decreased significantly.
\end{enumerate}

@@
